### 2021-05-02 doc/VerticalArrangements.tex
read through my sections and incoprated nits. Fixed some figure captions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title Page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Beyond ``Horizontal'' and ``Vertical'':\\ The Welfare Effects of Complex Integration\footnote{The analysis and conclusions set forth are those of the authors and do not indicate concurrence by other members of the Board research staff, by the Federal Reserve Board of Governors, by the Federal Trade Commission, or by its Commissioners.  Furthermore, the views expressed here should not be purported to reflect those of the U.S. Department of Justice.}}

@@

Returning to a merger between a preexisting integrated retailer/wholesaler $rw$ and a standalone retailer $s$, when the merged firm bargains with an unaffiliated wholesaler $v$ to supply retailer $r$, the first order condition becomes
\begin{equation}
\begin{split}
&[p^W_{rv} - c^W_{rv}]s_{rv} - \sum_{t \in \mathbb{R}^v \setminus \{r\}} [p^W_{tv} - c^W_{tv}] \Delta s_{tv}(\mathbb{W}^r \setminus \{v\}) = \\
&\frac{1-\lambda}{\lambda} \left([p_{rv} - p^W_{rv} - c^R_{rv}]s_{rv} - \sum_{x \in \mathbb{W}^r \setminus \{w,v\}} [p_{rx} - p^W_{rx} - c^R_{rx}] \Delta s_{rx}(\mathbb{W}^r \setminus \{v\})\right.\\
&- \underbrace{\sum_{t \in \{r,s\}}[p_{tw} - c^W_{tw} - c^R_{tw}] \Delta s_{tw}(\mathbb{W}^r \setminus \{v\})}_{\text{EDM recapture effect}}
- \underbrace{\sum_{x \in \mathbb{W}^s \setminus \{w\}} [p_{sx} - p^W_{sx} - c^R_{sx}] \Delta s_{sx}(\mathbb{W}^r \setminus \{v\})}_{\text{retail recapture leverage effect}}\\
&-\underbrace{\left.\sum_{t \in \mathbb{R}^w \setminus \{r,s\}} [p^W_{tw} - c^W_{tw}] \Delta s_{tw}(\mathbb{W}^r \setminus \{v\})\right)}_{\text{wholesale recapture leverage effect}}.
\end{split}
\label{eq: intdmerger ret upstream foc}
\end{equation}
Compared to equation \eqref{eq: vmerger ret upstream foc}, now the EDM recapture effect applies to sales of product $w$ at both affiliated retailers $r$ and $s$.  Furthermore, there is also the possibility that sales will be diverted to retailer $s$ should retailer $r$ lose access to product $v$, which creates an additional ``retail recapture leverage effect'' alongside the wholesale recapture effect.  Both of these additional terms tend to increase the bargaining leverage of the merged firm.  If we instead examined a merger between retailer/wholesaler $rw$ and an unaffiliated wholesaler, that would augment the EDM recapture effect with sales of retailer $r$ for two merged wholesalers, rather than sales through two merged retailers.  Similarly, the retail recapture leverage effect would be replaced with an additional wholesale recapture leverage effect for the newly merged wholesaler.  For a merger between two integrated retailer/wholesaler firms, all of these effects would enter.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Data Generating Process}
This section provides an overview of our methodology, with additional details appearing in the Appendix.  For vertical mergers, we only consider mergers between an unintegrated wholesaler and unintegrated retailer. In addition, we consider upstream and downstream mergers between either two unintegrated firms or between one vertically integrated and one unintegrated firm.  Finally, we assume that an integrated merger occurs between two vertically integrated firms\textbf{consider adding diagonal mergers by allowing for heterogeneity in the bargaining parameter.}.

We simulate markets by randomly sampling shares from a Dirichlet distribution for 2, 3, 4, or 5 retailers or wholesalers, respectively.\footnote{We parameterize the Dirichlet distribution so it is equivalent to a uniform distribution.  }  Because integrated mergers must have two vertically integrated incumbents in the pre-merger state, we increase the maximum allowable number of wholesalers and retailers in our simulated markets to 7. We also assume that in the pre-merger state, there are anywhere from 0 to 4 vertically integrated incumbents (6 for integrated mergers). We assume that vertically integrated firms are not siloed: integrated wholesalers supply inputs to non-integrated retailers and that vertically integrated retailers purchase inputs from non-integrated wholesalers. The price coefficient $\alpha$ is calibrated by assuming that in the pre-merger world, there is a vertically integrated outside option available to all customers.  The other goods are differenced relative to this option, which maintains the outside good normalization.  The market size is set to 1.

@@

We focus first on the results for consumers in the left-most panel of Figure \ref{fig:surplussum}.  The median change is negative, indicating harm, across all four types of mergers for both cost specifications.  However, the distributions and magnitudes differ.  In particular, there is only a partial rank-ordering of consumer harm across different types of mergers: consumer harm from integrated mergers first-order stochastically dominates consumer harm from all other merger  types, while downstream and upstream mergers each first-order stochastically dominate consumer harm from vertical mergers, but do not stochastically dominate one other.   Median consumer harm from integrated mergers is almost 14\% of pre-merger total expenditures, 1.25 times the magnitude of that from downstream mergers, 2.3 times the magnitude of that from upstream mergers, and more than 4 times the magnitude of vertical mergers.

@@

In terms of total welfare, approximately 18\% of vertical mergers and 15\% of integrated mergers with constant costs that are beneficial, whereas only 3\% of upstream mergers and 2\% of downstream mergers are beneficial. Moreover, there is a partial rank ordering of mergers, with total harm from downstream mergers first-order stochastically dominating total harm from upstream mergers, which dominates totals harm from vertical mergers. Median consumer harm from integrated mergers is about 9.6\% of pre-merger total expenditures, 1.5 times the magnitude of that from downstream mergers, 2.5 times the magnitude of that from upstream mergers, and more than 4.3 times the magnitude of vertical mergers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Vertically Integrated Incumbent Firms}
Here, we examine merger outcomes when a vertically integrated merges with either an unintegrated upstream or unintegrated downstream firm, or another integrated firm. We also examine merger outcomes for  mergers as the number of incumbent integrated non-merging rivals increases.

Figure \ref{fig:CVvertincumbBW_updownBW} depicts box and whisker plots summarizing consumer harm (top) and total harm (bottom) as the number of incumbent integrated firms increases from 0 to 6 firms. The box and whisker  plots when the number of  incumbent integrated firms equals 0 correspond to the results depicted in Figure 1 of \citeN{ST2020} and is included as reference.\footnote{An important difference between the simulations depicted in Figure \ref{fig:CVvertincumbBW} and those in Figure 1 of \citeN{ST2020} is that these results exclude downstream markets where prices are set according to a 2nd score auction}  For vertical mergers, the plots when the number of  incumbent integrated firms equals 1 summarize market outcomes where an unintegrated wholesaler merging with an unintegrated retailer, a single unintegrated 3rd party is integrated, and the remaining 3rd parties (if any) are unintegrated. By contrast,for downstream and upstream  mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, with no integrated third parties. Finally, for integrated mergers, the plots when the number of  incumbent integrated firms equals 2 depict a merger between two integrated firms, again with no integrated rivals.

Starting with vertical mergers, absent incumbent integration (0 integrated firms), about 36\% of vertical mergers benefit consumers and about 23\% of vertical mergers are net beneficial (i.e. increase total surplus).  The presence of a single 3rd party integrated firm (1 integrated firm) decreases the percentage of mergers that benefit consumers to 25\% and the percentage of mergers that increase  total surplus to 16\%, but leads median harm largely unchanged. Interestingly, while adding a second integrated incumbent to the market also lowers the percentage of beneficial mergers, adding a third integrated firm increases the percentage of mergers that benefits consumers to about 27\% but does not increase the percentage of net beneficial mergers. Adding a fourth integrated incumbent increases the percentage of mergers that benefit consumers to 32\% but decreases the percentage of net beneficial mergers to about 11\%. Taken together, Figure \ref{fig:CVvertincumbBW_updownBW} indicates that the presence of incumbent integrated firms is harmful, and that the harm does not meaningfully lessen as additional integrated incumbents are included.

@@

Figure \ref{fig:CVbargnobothBW}  depicts box and whisker plots summarizing the consumer (top) and total (bottom) welfare effects for downstream, upstream, and vertical mergers as the relative bargaining power parameter increases from 1/9 (retailers have the advantage to 9 (wholesalers have the advantage). also depicted are three sets of box and whisker plots that correspond to the number of incumbent integrated firms included in the simulated markets: 0 (light blue), 1 (darker blue), and 4 (darkest blue)incumbent vertically integrated firms.  As in Figure \ref{fig:CVvertincumbBW_updownBW}, plots when the number of integrated firms is 0 assume that pre-merger, no firms in the market are vertically integrated: these are comparable to those in \citeN{ST2020}. For vertical mergers (left), the plots when the number of  incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when a single 3rd party is integrated. By contrast, for upstream (middle) and downstream (left) mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, and all 3rd parties are unintegrated. Finally, plots with 4 integrated firms increases the number of rival incumbent integrated firms in upstream and downstream mergers to 3 and in vertical mergers to 4 .

For vertical mergers, the presence of integrated incumbents has little impact on the strong, positive relationship between bargaining power and harm for vertical mergers. Figure \ref{fig:CVbargnobothBW} shows that consumer harm when there are fewer incumbent integrated firms first-order stochastically dominates consumer harm when there are more incumbent integrated firms, the difference is smallest when wholesalers and retailers have equal bargaining power and greatest as wholesaler power increases relative to retailer power. Across all three incumbent integrated firm scenarios, median consumer harm decreases from about 7\% when retailers have the advantage to -24\% when wholesalers have the advantage. Similarly, median net harm decreases from about 2.9\% of pre-merger expenditures when retailers have relatively more bargaining power to about -14\% of pre-merger expenditures when wholesalers have relatively more bargaining power.

Turning to upstream mergers, when no incumbent firms are in the market (0 integrated firms), there is a positive relationship between bargaining power and both consumer and total harm.  Median consumer harm increases from 2.2\% of pre-merger expenditures when retailers have relatively more bargaining power to about 29\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from 0.7\% of pre-merger expenditures when retailers have relatively more bargaining power to about 12\% when wholesalers have relatively more bargaining power. By contrast, for mergers between an integrated and unintegrated upstream supplier (1 integrated firms), there is a positive relationship between bargaining power and both consumer and total surplus. Median consumer harm decreases from about 8.4\% of pre-merger expenditures when retailers have relatively more bargaining power to about -24\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from about 3.5\% of pre-merger expenditures when retailers have relatively more bargaining power to -8.4\% when wholesalers have relatively more bargaining power. Adding 2 additional incumbent integrated firms (4 integrated firms) mitigates some of the adverse effects of mergers while still preserving the positive relationship between bargaining power and harm.

@@

\textbf{Describe the counterintuitive result that increasing retailer bargaining power harms retailers but help wholesalers.}

@@

This paper relaxes two strong assumptions made in \citeN{ST2020}: that mergers occur in vertical supply chains where in the pre-merger state, no firms are vertically integrated, and that all firms employ constant marginal cost technology. Relaxing these constraints is important as competition authorities are called upon to investigate mergers where either the merging parties or a third party are already vertically integrated, or where the presence of a scarce input could reduce the benefit from EDM or increase the harm from RRC. Here, we show using both numerical simulations that relaxing these assumptions can result in merger outcomes that are markedly different from those when these assumptions are maintained. We also show using simulation that relative bargaining power can serve as a useful indicia for predicting consumer and total harm from mergers where these two assumptions are relaxed. Finally, we intend to use two recent trash mergers to investigate the importance of these assumptions in predicting merger harm. Our hope is that future research will focus on some of the other strong assumptions made in \citeN{ST2020}, such as how the presence of either two-part tariffs or large fixed costs affect the Nash bargaining game and therefore merger outcomes.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/surplussum.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer, retailer, wholesaler, and total surplus. Each blue box depicts the effects assuming that firms face constant marginal costs, while each orange box depicts the effects assuming that firms face linear marginal costs. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:surplussum}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer (blue,left) and total (orange,right) surplus as the number of vertically integrated firms present in a market change.  Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/surplussum_cost.png}
\caption{The figure displays box and whisker plots summarizing the extent to which merger outcomes change according to 4 different cost scenarios. The boxes in the top row (``All") either assume all firms face either constant marginal costs (blue,left) or linear marginal costs (orange, right). The boxes in the bottom row (``Party") eitehr assume that the merging parties face constant marginal costs while other firms face linear marginal costs (blue, left), or the merging parties face linear marginal costs while other firms face constant marginal costs (orange, right). Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:surplussumcost}
\end{sidewaysfigure}
\end{document}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title Page
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Beyond ``Horizontal'' and ``Vertical'':\\ The Welfare Effects of Complex Integration\footnote{The analysis and conclusions set forth are those of the authors and do not indicate concurrence by other members of the Board research staff, by the Federal Reserve Board of Governors, or by the Federal Trade Commission.  Furthermore, the views expressed here should not be purported to reflect those of the U.S. Department of Justice.}}

@@

Returning to a merger between a preexisting integrated retailer/wholesaler $rw$ and a standalone retailer $s$, when the merged firm bargains with an unaffiliated wholesaler $v$ to supply retailer $r$, the first order condition becomes
\begin{equation}
\begin{split}
&[p^W_{rv} - c^W_{rv}]s_{rv} - \sum_{t \in \mathbb{R}^v \setminus \{r\}} [p^W_{tv} - c^W_{tv}] \Delta s_{tv}(\mathbb{W}^r \setminus \{v\}) = \\
&\frac{1-\lambda}{\lambda} \left([p_{rv} - p^W_{rv} - c^R_{rv}]s_{rv} - \sum_{x \in \mathbb{W}^r \setminus \{w,v\}} [p_{rx} - p^W_{rx} - c^R_{rx}] \Delta s_{rx}(\mathbb{W}^r \setminus \{v\})\right.\\
&- \underbrace{\sum_{t \in \{r,s\}}[p_{tw} - c^W_{tw} - c^R_{tw}] \Delta s_{tw}(\mathbb{W}^r \setminus \{v\})}_{\text{EDM recapture effect}}
- \underbrace{\sum_{x \in \mathbb{W}^s \setminus \{w\}} [p_{sx} - p^W_{sx} - c^R_{sx}] \Delta s_{sx}(\mathbb{W}^r \setminus \{v\})}_{\text{retail recapture leverage effect}}\\
&-\underbrace{\left.\sum_{t \in \mathbb{R}^w \setminus \{r,s\}} [p^W_{tw} - c^W_{tw}] \Delta s_{tw}(\mathbb{W}^r \setminus \{v\})\right)}_{\text{wholesale recapture leverage effect}}.
\end{split}
\label{eq: intdmerger ret upstream foc}
\end{equation}
Compared to equation \eqref{eq: vmerger ret upstream foc}, now the EDM recapture effect applies to sales of product $w$ at both affiliated retailers $r$ and $s$.  Furthermore, there is also the possibility that sales will be diverted to retailer $s$ should retailer $r$ lose access to product $v$, which creates an additional ``retail recapture leverage effect'' alongside the wholesale recapture effect.  Both of these additional terms tend to increase the bargaining leverage of the merged firm.  If we instead examined a merger between retailer/wholesaler $rw$ and an unaffiliated wholesaler, that would augment the EDM recapture effect with sales of retailer $r$ for two merged wholesalers, rather than sales through two merged retailers.  Similarly, the retail recapture leverage effect would be replaced with an additional wholesale recapture leverage effect for the newly merged wholesaler.  For a merger between two integrated retailer/wholesaler firms, all of these effects would enter.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Data Generating Process}
This section provides an overview of our methodology, with additional details appearing in the Appendix.  For vertical mergers, we only consider mergers between an unintegrated wholesaler and unintegrated retailer. In addition, we consider upstream and downstream mergers between either two unintegrated firms or between one vertically integrated and one unintegrated firm.  Finally, we assume that an integrated merger occurs between two vertically integrated firms

%\textbf{consider adding diagonal mergers by allowing for heterogeneity in the bargaining parameter.}.

We simulate markets by randomly sampling shares from a Dirichlet distribution for 2, 3, 4, or 5 retailers or wholesalers, respectively.\footnote{We parameterize the Dirichlet distribution so it is equivalent to a uniform distribution.  }  Because integrated mergers must have two vertically integrated incumbents in the pre-merger state, for these simulations we increase the maximum allowable number of wholesalers and retailers in our simulated markets to 7. We also assume that in the pre-merger state, there are anywhere from 0 to 4 vertically integrated incumbents (6 for integrated mergers). We assume that vertically integrated firms are not siloed: integrated wholesalers supply inputs to non-integrated retailers and that vertically integrated retailers purchase inputs from non-integrated wholesalers. The price coefficient $\alpha$ is calibrated by assuming that in the pre-merger world, there is a vertically integrated outside option available to all customers.  The other goods are differenced relative to this option, which maintains the outside good normalization.  The market size is set to 1.

@@

We focus first on the results for consumers in the left-most panel of Figure \ref{fig:surplussum}.  The median change is negative, indicating harm, across all four types of mergers.  However, the distributions and magnitudes differ.  In particular, there is only a partial rank-ordering of consumer harm across different types of mergers: consumer harm from integrated mergers first-order stochastically dominates consumer harm from all other merger  types, while downstream and upstream mergers each first-order stochastically dominate consumer harm from vertical mergers, but do not stochastically dominate one other.   Median consumer harm from integrated mergers is almost 14\% of pre-merger total expenditures, 1.25 times the magnitude of that from downstream mergers, 2.3 times the magnitude of that from upstream mergers, and more than 4 times the magnitude of vertical mergers.

@@

In terms of total welfare, approximately 18\% of vertical mergers and 15\% of integrated mergers are beneficial, whereas only 3\% of upstream mergers and 2\% of downstream mergers are beneficial. Moreover, there is a partial rank ordering of mergers, with total harm from downstream mergers first-order stochastically dominating total harm from upstream mergers, which dominates totals harm from vertical mergers. Median consumer harm from integrated mergers is about 9.6\% of pre-merger total expenditures, 1.5 times the magnitude of that from downstream mergers, 2.5 times the magnitude of that from upstream mergers, and more than 4.3 times the magnitude of vertical mergers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Vertically Integrated Incumbent Firms}
Here, we examine merger outcomes when a vertically integrated firm merges with either an unintegrated upstream or unintegrated downstream firm, or another integrated firm. We also examine merger outcomes for  mergers as the number of incumbent integrated non-merging rivals increases.

Figure \ref{fig:CVvertincumbBW} depicts box and whisker plots summarizing consumer harm (top) and total harm (bottom) as the number of incumbent integrated firms increases from 0 to 6 firms. The box and whisker  plots when the number of  incumbent integrated firms equals 0 correspond to the results depicted in Figure 1 of \citeN{ST2020} and is included as reference.\footnote{An important difference between the simulations depicted in Figure \ref{fig:CVvertincumbBW} and those in Figure 1 of \citeN{ST2020} is that these results exclude downstream markets where prices are set according to a 2nd score auction. \citeN{ST2020} also do not consider mergers between integrated firms. }  For vertical mergers, the plots when the number of  incumbent integrated firms equals 1 summarize market outcomes where an unintegrated wholesaler merges with an unintegrated retailer, a single unintegrated 3rd party is integrated, and the remaining 3rd parties (if any) are unintegrated. By contrast, for downstream and upstream  mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, with no integrated third parties. Finally, for integrated mergers, the plots when the number of  incumbent integrated firms equals 2 depict a merger between two integrated firms, again with no integrated rivals.

Starting with vertical mergers, absent incumbent integration (0 integrated firms), about 36\% of vertical mergers benefit consumers and about 23\% of vertical mergers are net beneficial (i.e. increase total surplus).  The presence of a single 3rd party integrated firm (1 integrated firm) decreases the percentage of mergers that benefit consumers to 25\% and the percentage of mergers that increase  total surplus to 16\%, but leads median harm largely unchanged. Interestingly, while adding a second integrated incumbent to the market also lowers the percentage of beneficial mergers, adding a third integrated firm increases the percentage of mergers that benefit consumers to about 27\% but does not increase the percentage of net beneficial mergers. Adding a fourth integrated incumbent further increases the percentage of mergers that benefit consumers to 32\% but decreases the percentage of net beneficial mergers to about 11\%. Taken together, Figure \ref{fig:CVvertincumbBW} indicates that the presence of incumbent integrated firms is harmful, and that the harm does not meaningfully lessen as additional integrated incumbents are included.

@@

Figure \ref{fig:CVbargnobothBW}  depicts box and whisker plots summarizing the consumer (top) and total (bottom) welfare effects for downstream, upstream, and vertical mergers as the relative bargaining power parameter increases from 1/9 (retailers have the advantage to 9 (wholesalers have the advantage). also depicted are three sets of box and whisker plots that correspond to the number of incumbent integrated firms included in the simulated markets: 0 (light blue), 1 (darker blue), and 4 (darkest blue) incumbent vertically integrated firms.  As in Figure \ref{fig:CVvertincumbBW}, plots when the number of integrated firms is 0 assume that pre-merger, no firms in the market are vertically integrated: these are comparable to those in \citeN{ST2020}. For vertical mergers (left), the plots when the number of  incumbent integrated firms equals 1 depict the outcome from an unintegrated wholesaler merging with an unintegrated retailer when a single 3rd party is integrated. By contrast, for upstream (middle) and downstream (left) mergers, the plots when the number of  incumbent integrated firms equals 1 depict the outcome of a merger between an integrated firm and an unintegrated rival, and all 3rd parties are unintegrated. Finally, plots with 4 integrated firms increases the number of rival incumbent integrated firms in upstream and downstream mergers to 3 and in vertical mergers to 4 .

For vertical mergers, the presence of integrated incumbents has little impact on the strong, positive relationship between bargaining power and harm for vertical mergers. Figure \ref{fig:CVbargnobothBW} shows that consumer harm when there are fewer incumbent integrated firms first-order stochastically dominates consumer harm when there are more incumbent integrated firms, the difference is smallest when wholesalers and retailers have equal bargaining power and greatest as wholesaler power increases relative to retailer power. Across all three incumbent integrated firm scenarios, median consumer harm decreases from about 7\% when retailers have the advantage to -24\% when wholesalers have the advantage. Similarly, median total harm decreases from about 2.9\% of pre-merger expenditures when retailers have relatively more bargaining power to about -14\% of pre-merger expenditures when wholesalers have relatively more bargaining power.

Turning to upstream mergers, when no incumbent firms are in the market (0 integrated firms), there is a positive relationship between bargaining power and both consumer and total harm.  Median consumer harm increases from 2.2\% of pre-merger expenditures when retailers have relatively more bargaining power to about 29\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from 0.7\% of pre-merger expenditures when retailers have relatively more bargaining power to about 12\% when wholesalers have relatively more bargaining power. By contrast, for mergers between an integrated and unintegrated upstream supplier (1 integrated firm), there is a positive relationship between bargaining power and both consumer and total surplus. Median consumer harm decreases from about 8.4\% of pre-merger expenditures when retailers have relatively more bargaining power to about -24\% of pre-merger expenditures when wholesalers have relatively more bargaining power. Likewise, median total harm increases from about 3.5\% of pre-merger expenditures when retailers have relatively more bargaining power to -8.4\% when wholesalers have relatively more bargaining power. Adding 2 additional incumbent integrated firms (4 integrated firms) mitigates some of the adverse effects of mergers while still preserving the positive relationship between bargaining power and harm.

@@

%\textbf{Describe the counterintuitive result that increasing retailer bargaining power harms retailers but help wholesalers.}

@@

This paper relaxes two strong assumptions made in \citeN{ST2020}: that mergers occur in vertical supply chains where in the pre-merger state, no firms are vertically integrated, and that all firms employ constant marginal cost technology. Relaxing these constraints is important as competition authorities are called upon to investigate mergers where either the merging parties or a third party are already vertically integrated, or where the presence of a scarce input could reduce the benefit from EDM or increase the harm from RRC. Here, we show using numerical simulations that relaxing these assumptions can result in merger outcomes that are markedly different from those when these assumptions are maintained. We also show using simulation that relative bargaining power can serve as a useful indicia for predicting consumer and total harm from mergers where these two assumptions are relaxed. Finally, we intend to use two recent trash mergers to investigate the importance of these assumptions in predicting merger harm. Our hope is that future research will focus on some of the other strong assumptions made in \citeN{ST2020}, such as: how the presence of siloed firms, two-part tariffs, or large fixed costs affect the Nash bargaining game and therefore merger outcomes.

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/surplussum.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer, retailer, wholesaler, and total surplus. Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:surplussum}
\end{sidewaysfigure}

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/CVvertincumbBW.png}
\caption{The figure displays box and whisker plots summarizing the extent to which mergers affect consumer (top row) and total (bottom) surplus as the number of vertically integrated firms present in a market change.  Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:CVvertincumbBW}
\end{sidewaysfigure}

@@

\begin{sidewaysfigure}
\centering
\includegraphics[scale=0.9]{../output/surplussum_cost.png}
\caption{The figure displays box and whisker plots summarizing the extent to which merger outcomes change according to 4 different cost scenarios. The boxes in the top row (``All") either assume all firms face either constant marginal costs (blue,left) or linear marginal costs (orange, right). The boxes in the bottom row (``Party") either assume that the merging parties face constant marginal costs while other firms face linear marginal costs (blue, left), or the merging parties face linear marginal costs while other firms face constant marginal costs (orange, right). Whiskers depict the $5^{th}$ and $95^{th}$ percentiles of a particular outcome, boxes depict the $25^{th}$ and $75^{th}$ percentiles, and the solid horizontal line depicts the median. }
\label{fig:surplussumcost}
\end{sidewaysfigure}
\end{document}
